Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2024-01-30 paper-correlations.tex
Regroup images

\begin{figure*}[tp]
  \centering
  \subfigure[Lineal-path correlation function]{
    \includegraphics[width=0.3\linewidth]{images/l2.png}
    \label{fig:l2}}
  \hfill
  \subfigure[Two-point correlation function]{
    \includegraphics[width=0.3\linewidth]{images/s2.png}
    \label{fig:s2}}
  \hfill
  \subfigure[Surface-surface correlation function]{
    \includegraphics[width=0.3\linewidth]{images/ss.png}
    \label{fig:ss}}
  \vskip\baselineskip
  \subfigure[Surface-void correlation function]{
    \includegraphics[width=0.3\linewidth]{images/sv.png}
    \label{fig:sv}}
  \hfill
  \subfigure[Pore size correlation function]{
    \includegraphics[width=0.3\linewidth]{images/ps.png}
    \label{fig:ps}}
  \hfill
  \subfigure[Chord length correlation function]{
    \includegraphics[width=0.3\linewidth]{images/cl.png}
    \label{fig:cl}}
  \caption[]{A comparison of calculated and theoretical values of correlation
    functions for overlapping disks and balls.}
  \label{fig:verification}
\end{figure*}

@@

At first, we compare the three ceramic images which have somewhat different
genesis depending on the mixture ratio of powder and binder used for firing the
material \cite{ceramics}. As observed from \cref{fig:real-xct}, the structure of
all three samples is visually very similar. The lineal path function supports
this observation. The $S_2$ functions are also quite similar except for Sample 2
having higher porosity that results in elevated curve (the tail for two-point
probability as expected fluctuates around porosity squared). Notably, $S_2$ and
$C_2$ are almost identical -- this indicates that pore phase represents a single
large cluster. Recall that for $S_2$, $C_2$ and $L_2$ functions the value at
correlation length of 0 is the porosity value. Surface functions, on the other
hand, feel significant difference between samples in how pore-solid interface is
organized. This is further evidenced on pore-size and chord-length graphs, with
Ceramic 2 sample with the highest porosity having larger amount of large
pores. Both $P$ and $p$ show very similar behavior, but for different
correlation lengths; such behavior is attributed to quite regular pores shapes
in all ceramics samples -- this is evident from much larger differences between
pore-size and chord-length for SEM images considered next.

\begin{figure*}[pt]
  \centering
  \subfigure[Ceramic 1]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic1.png}}
  \hfill
  \subfigure[Ceramic 2]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic2.png}}
  \hfill
  \subfigure[Ceramic 3]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic3.png}}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/s2-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/c2-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/l2-xct.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/surf2-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/surfvoid-xct.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/pore-size-xct.png}
  \includegraphics[width=0.32\linewidth]{samples/chord-length-xct.png}
  \caption[]{Visualization of XCT images and plots of correlation functions for
    these images.}
  \label{fig:real-xct}
\end{figure*}

The \cref{fig:real-sem} presents three 2D images of sandstones and carbonate
with computed correlation functions. In this example, unlike for quite isotropic
ceramics, we evaluated CFs in the directional manner. The $S_2$ graph
immediately provides us with porosity values. It also highlights severe
anisotropy in sandstone SEM images. For 2D images numerous patches of pores are
not connected to each other and, thus, $S_2$ and $C_2$ are very
different. Cluster functions show how connectivity decays with length, also we
can easily observe if given pore structure percolates in a given direction -
none of the images percolates (we highlighted the clusters to make this
statement easily verifiable with the naked eye). Note that $C_2$ and $L_2$ were
computed with correlations lengths equal to image dimensions explore
connectivity -- for other CFs this leads to noised curves due to smaller number
of sampled points (for example, see correlation functions' graphs in
\cite{Euras2012}).  Lineal paths are close for sandstones (with lower $L_2$
values for Sandstone 1 where large pores are occluded with clays or grain
fragments), but for carbonate with much smaller apertures of pores it is way
below $L_2$ for sandstones. Also, $L_2$ suggests that computations with
correlations lengths beyond where lineal path goes to zero are not very
informative. For very irregular pores on 2D SEM images, $P$ and $p$ functions
look much less similar compared to the ceramics case on \cref{fig:real-xct}. For
2D images it is also visually easy to analyze full CF maps, as presented on
\cref{fig:real-sem-map} for Sandstone 1 sample. Full correlation maps show
expanded version of directional functions -- asymmetry is immediately evidenced
from $S_2$, $F_{ss}$ and $F_{sv}$ maps. Moreover, the map for $C_2$ shows higher
connectivity in y-direction as compared to x-axis. The interpretation of full
correlation maps is less straightforward, but anisotropy patterns in directions
different from those presented in \cref{fig:directions} can be spotted.

\begin{figure*}[pt]
  \centering
  \subfigure[Sandstone 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone1.png}}
  \hfill
  \subfigure[Sandstone 2]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone2.png}}
  \hfill
  \subfigure[Carbonate 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/carbonate1.png}}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/s2-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/c2-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/l2-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/surf2-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/surfvoid-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.32\linewidth]{samples/pore-size-sem.png}
  \includegraphics[width=0.32\linewidth]{samples/chord-length-sem.png}
  \caption[]{SEM images and plots of correlation functions for these
    images. Different clusters of the void phase are shown in color.}
  \label{fig:real-sem}
\end{figure*}

>>>>

\begin{figure*}[tp]
  \centering
  \subfigure[Lineal-path correlation function]{
    \includegraphics[width=0.45\linewidth]{images/l2.png}
    \label{fig:l2}}
  \hfill
  \subfigure[Two-point correlation function]{
    \includegraphics[width=0.45\linewidth]{images/s2.png}
    \label{fig:s2}}
  \vskip\baselineskip
  \subfigure[Surface-surface correlation function]{
    \includegraphics[width=0.45\linewidth]{images/ss.png}
    \label{fig:ss}}
  \hfill
  \subfigure[Surface-void correlation function]{
    \includegraphics[width=0.45\linewidth]{images/sv.png}
    \label{fig:sv}}
  \vskip\baselineskip
  \subfigure[Pore size correlation function]{
    \includegraphics[width=0.45\linewidth]{images/ps.png}
    \label{fig:ps}}
  \hfill
  \subfigure[Chord length correlation function]{
    \includegraphics[width=0.45\linewidth]{images/cl.png}
    \label{fig:cl}}
  \caption[]{A comparison of calculated and theoretical values of correlation
    functions for overlapping disks and balls.}
  \label{fig:verification}
\end{figure*}

@@

At first, we compare the three ceramic images (\cref{fig:real-data}, at the top)
which have somewhat different genesis depending on the mixture ratio of powder
and binder used for firing the material \cite{ceramics}. As observed from
\cref{fig:real-xct-cfs}, the structure of all three samples is visually very
similar. The lineal path function supports this observation. The $S_2$ functions
are also quite similar except for Sample 2 having higher porosity that results
in elevated curve (the tail for two-point probability as expected fluctuates
around porosity squared). Notably, $S_2$ and $C_2$ are almost identical -- this
indicates that pore phase represents a single large cluster. Recall that for
$S_2$, $C_2$ and $L_2$ functions the value at correlation length of 0 is the
porosity value. Surface functions, on the other hand, feel significant
difference between samples in how pore-solid interface is organized. This is
further evidenced on pore-size and chord-length graphs, with Ceramic 2 sample
with the highest porosity having larger amount of large pores. Both $P$ and $p$
show very similar behavior, but for different correlation lengths; such behavior
is attributed to quite regular pores shapes in all ceramics samples -- this is
evident from much larger differences between pore-size and chord-length for SEM
images considered next.

\begin{figure*}[pt]
  \centering
  \subfigure[Ceramic 1]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic1.png}}
  \hfill
  \subfigure[Ceramic 2]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic2.png}}
  \hfill
  \subfigure[Ceramic 3]{
    \includegraphics[width=0.3\linewidth]{samples/Ceramic3.png}}
  \vskip\baselineskip
  \subfigure[Sandstone 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone1.png}}
  \hfill
  \subfigure[Sandstone 2]{
    \includegraphics[width=0.3\linewidth, frame]{samples/sandstone2.png}}
  \hfill
  \subfigure[Carbonate 1]{
    \includegraphics[width=0.3\linewidth, frame]{samples/carbonate1.png}}
  \caption[]{Visualization of XCT (top) and SEM (bottom) images used for
    computation of correlation functions. Different clusters of the void phase
    on SEM samples are shown in color.}
  \label{fig:real-data}
\end{figure*}
\begin{figure*}[pt]
  \centering
  \includegraphics[width=0.45\linewidth]{samples/s2-xct.png}
  \hfill
  \includegraphics[width=0.45\linewidth]{samples/c2-xct.png}
  \vskip\baselineskip
  \includegraphics[width=0.45\linewidth]{samples/l2-xct.png}
  \vskip\baselineskip
  \includegraphics[width=0.45\linewidth]{samples/surf2-xct.png}
  \hfill
  \includegraphics[width=0.45\linewidth]{samples/surfvoid-xct.png}
  \vskip\baselineskip
  \includegraphics[width=0.45\linewidth]{samples/pore-size-xct.png}
  \hfill
  \includegraphics[width=0.45\linewidth]{samples/chord-length-xct.png}
  \caption[]{Plots of correlation functions for XCT samples on
    \cref{fig:real-data}.}
  \label{fig:real-xct-cfs}
\end{figure*}

The \cref{fig:real-data} (at the bottom) presents three 2D images of sandstones
and carbonate with computed correlation functions at \cref{fig:real-sem-cfs}. In
this example, unlike for quite isotropic ceramics, we evaluated CFs in the
directional manner. The $S_2$ graph immediately provides us with porosity
values. It also highlights severe anisotropy in sandstone SEM images. For 2D
images numerous patches of pores are not connected to each other and, thus,
$S_2$ and $C_2$ are very different. Cluster functions show how connectivity
decays with length, also we can easily observe if given pore structure
percolates in a given direction - none of the images percolates (we highlighted
the clusters to make this statement easily verifiable with the naked eye). Note
that $C_2$ and $L_2$ were computed with correlations lengths equal to image
dimensions explore connectivity -- for other CFs this leads to noised curves due
to smaller number of sampled points (for example, see correlation functions'
graphs in \cite{Euras2012}).  Lineal paths are close for sandstones (with lower
$L_2$ values for Sandstone 1 where large pores are occluded with clays or grain
fragments), but for carbonate with much smaller apertures of pores it is way
below $L_2$ for sandstones. Also, $L_2$ suggests that computations with
correlations lengths beyond where lineal path goes to zero are not very
informative. For very irregular pores on 2D SEM images, $P$ and $p$ functions
look much less similar compared to the ceramics case on
\cref{fig:real-xct-cfs}. For 2D images it is also visually easy to analyze full
CF maps, as presented on \cref{fig:real-sem-map} for Sandstone 1 sample. Full
correlation maps show expanded version of directional functions -- asymmetry is
immediately evidenced from $S_2$, $F_{ss}$ and $F_{sv}$ maps. Moreover, the map
for $C_2$ shows higher connectivity in y-direction as compared to x-axis. The
interpretation of full correlation maps is less straightforward, but anisotropy
patterns in directions different from those presented in \cref{fig:directions}
can be spotted.
\begin{figure*}[pt]
  \centering
  \includegraphics[width=0.45\linewidth]{samples/s2-sem.png}
  \hfill
  \includegraphics[width=0.45\linewidth]{samples/c2-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.45\linewidth]{samples/l2-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.45\linewidth]{samples/surf2-sem.png}
  \hfill
  \includegraphics[width=0.45\linewidth]{samples/surfvoid-sem.png}
  \vskip\baselineskip
  \includegraphics[width=0.45\linewidth]{samples/pore-size-sem.png}
  \hfill
  \includegraphics[width=0.45\linewidth]{samples/chord-length-sem.png}
  \caption[]{Plots of correlation functions for SEM samples on
    \cref{fig:real-data}}
  \label{fig:real-sem-cfs}
\end{figure*}
